\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}
{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}
\begin{document}
\begin{titlepage} % Suppresses displaying the page number on the title page and the subsequent page counts as page 1

	\raggedleft\rule{1pt}{\textheight} % Vertical line
	\hspace{0.05\textwidth} % Whitespace between the vertical line and title page text
	\parbox[b]{0.75\textwidth}
    { % Paragraph box for holding the title page text, adjust the width to move the title page left or right on the page
		
		{\Huge\bfseries MIT World Peace University \\[0.5\baselineskip] \ Python Programming}\\[2\baselineskip] % Title
		{\large\textit{Assignment 1}}\\[4\baselineskip] % Subtitle or further description
		{\Large\textsc{Naman Soni Roll No. 10}} % Author name, lower case for consistent small caps
		
		\vspace{0.5\textheight} % Whitespace between the title block and the publisher
	}

\end{titlepage}
\tableofcontents
\pagebreak
\section{\textbf{Problem Statement}}
Introduction to basic Python Commands.
\section{\textbf{Aim}}
To learn the basics of the python programming language and understand fundamental syntax and semantics of Python Programming.
\section{\textbf{Objectives}}
\begin{enumerate}
    \item To learn the basics of python programming language.
    \item To learn the variable declaration,user input and output of the programming language.
\end{enumerate}
\section{\textbf{Theory}}
\subsection{\textbf{\textit{Introduction to Python}}}
Python is a high-level, interpreted programming language that was first released in 1991. It is known for its simple and easy-to-read syntax, making it a popular choice for beginners and experienced developers alike. Python can be used for a wide range of tasks, including web development, scientific computing, data analysis, artificial intelligence, and more. It is an open-source language with a large and supportive community, and has a vast library of modules and tools available to users.
\subsection{\textbf{\textit{Basic Commands in Python}}}
\begin{enumerate}
    \item Print: display a message on the screen. For example, print (``Hello World '').
    \item Variable assignment: assign a value to a variable. For example, x = 5.
    \item Arithmetic operations: perform mathematical operations such as addition, subtraction, multiplication, and division. For example, x + 5, x - 5, x * 5, and x / 5
    \item Conditional statements: make decisions based on conditions. For example, if x > 5: print (``x is greater than 5'').
    \item Loops: repeat a block of code multiple times. For example, for i in range (5): print (i).
    \item Importing modules: use pre-existing code from other sources. For example, import math to use mathematical functions.
\end{enumerate}
\subsection{\textbf{\textit{Standard Data Types}}}
The data stored in memory can be of many types. Python has various standard data types
that are used to define the operations possible on them and the storage method for each of
them.\\
Python has 5 standard data types:
\begin{itemize}
    \item Numbers
    \item String
    \item List 
    \item Tuple 
    \item Dictionary
\end{itemize}
\section{\textbf{Platform}}
Python\\
Mac OS 64-bit\\
Visual Studio Code
\section{\textbf{Code Input/Output}}
\begin{lstlisting}[language=python,caption=print function input]
    print("Hello world!")
\end{lstlisting}
\begin{lstlisting}
    Hello world!
\end{lstlisting}
\begin{lstlisting}[language=python,caption=declaring variable input]
# declare variable
a = 3
print(a)
print(type(a))
b = 3.5
print(type(b))
print(b)
c = "Hello"
print(c)
print(type(c))
\end{lstlisting}
\begin{lstlisting}
3
<class 'int'>
<class 'float'>
3.5
Hello
<class 'str'>
\end{lstlisting}
\begin{lstlisting}[language=python,caption=performing arithmetic operation input]
a = 5
b = 6
print(a + b)
\end{lstlisting}
\begin{lstlisting}
11
\end{lstlisting}
\begin{lstlisting}[language=python]
a = input("Enter First Name:")
b = input("Enter Last Name:")
print(a + b)
\end{lstlisting}
\begin{lstlisting}
naman soni 
\end{lstlisting}
\begin{lstlisting}[language=python]
x = int(input("Enter the first number:"))
y = int(input("Enter the second number:"))
z = x + y
print("Sum is:", z)
\end{lstlisting}
\begin{lstlisting}
Sum is: 66
\end{lstlisting}
\begin{lstlisting}[language=python]
x = float(input("Enter the first number:"))
y = float(input("Enter the second number:"))
z = x + y
print("Sum is:", z)
\end{lstlisting}
\begin{lstlisting}
Sum is: 110.6943    
\end{lstlisting}
\begin{lstlisting}[language=python]
a = 45
b = 65
if (a == b):
    print("equal")
else:
    print("not equal")
\end{lstlisting}
\begin{lstlisting}
not equal
\end{lstlisting}
\begin{lstlisting}[language=python]
my_list = [12, 13, 56, 43]
print(my_list)
\end{lstlisting}
\begin{lstlisting}
[12, 13, 56, 43]
\end{lstlisting}
\section{\textbf{Conclusion}}
Understood the basic commands, statements and syntax of python programming language.
\section{\textbf{FAQ's}}
\begin{enumerate}
	\item What are the features of python language?\\
	\begin{itemize}
		\item 1. Easy to Learn: Python has a very simple  and elegant syntadx which is much easier to understand then other programming languages.
		\item 2. Interpreted  Language:  Python is an interpreted language which means thet the written code is not actually translated to a computer-readable format at the time of execution.
		\item 3. Cross-Platform Language: Python can run equally on diffrent platforms such as Windows, Linux, Unix and Macintosh, etc. Thus, we can develop applications with more compatibility with others.
		\item 4. Object-Oriented Language: Python supports object-oriented language and concepts of classses and objects come into existence.
		\item 5. Dynamic Typing: Python uses dynamic typing, which means that variables can change type during runtime, making it more flexible than statically typed languages.
		\item 6. Large Standard Library: Python has a large standard library that includes modules for various tasks, such as file I/O, regular expressions, and Internet protocols.
	\end{itemize}
	\item Explain print () statement of python language.\\
	\textbf{Ans.} The print() function in Python is used to display text or output data on the screen. It can be used to print simple text strings, variables, expressions, or complex objects such as lists, dictionaries, and objects.\\
	The basic syntax of print() function is: 
	\begin{lstlisting}[language=python, caption=Basic Syntax of Print function]
		print(object(s), sep=separator, end=ending, file=file, flush=flush)
	\end{lstlisting}
	\item How to write single line and multi-line comments in python programming language.\\
	\textbf{Ans.} In Python, you can add comments to your code to make it more readable and explain what it does. There are two ways to write comments in Python:
		\begin{enumerate}
			\item Single-line comments: Single-line comments start with a (hash) symbol and continue to the end of the line. They are used to add brief explanations or notes to your code.
			\begin{lstlisting}
				# This is a single line comment
			\end{lstlisting}
			\item Multi-line comments: Multi-line comments are used to add longer explanations or notes to your code. They start with (''') or (""") and end with the same sequence.
			\begin{lstlisting}
				'''
				This is a 
				multiline 
				comment
				'''
				
				OR 
				
				"""
				This is also 
				a multline 
				comment
				"""
			\end{lstlisting}
		\end{enumerate}
\end{enumerate}
\end{document}
