\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{graphicx}

\begin{document}
\begin{titlepage} % Suppresses displaying the page number on the title page and the subsequent page counts as page 1
	
	\raggedleft\rule{1pt}{\textheight} % Vertical line
	\hspace{0.05\textwidth} % Whitespace between the vertical line and title page text
	\parbox[b]{0.75\textwidth}
	{ % Paragraph box for holding the title page text, adjust the width to move the title page left or right on the page
		
		{\Huge\bfseries MIT World Peace University \\[0.5\baselineskip] \ Peace}\\[2\baselineskip] % Title
		{\large\textit{Film Review}}\\[4\baselineskip] % Subtitle or further description
		{\Large\textsc{Naman Soni Roll No. 10}} % Author name, lower case for consistent small caps
		
		\vspace{0.5\textheight} % Whitespace between the title block and the publisher
	}
	
\end{titlepage}
\tableofcontents
\pagebreak
\begin{center}
    \textbf{Invictus}
    \textbf{Film Review}
\end{center}
``Invictus'' is a powerful and inspiring film that explores themes of leadership, unity, and reconciliation. Directed by Clint Eastwood, the film is based on the true story of Nelson Mandela's efforts to unite a deeply divided South Africa through the country's rugby team.

The film's main character is Nelson Mandela, played brilliantly by Morgan Freeman. Freeman captures Mandela's quiet strength and determination, as well as his compassion and ability to inspire others. Matt Damon also delivers a strong performance as Francois Pienaar, the captain of the South African rugby team.

The social impact of the film is significant. ``Invictus'' shows how sport can be a powerful tool for social change, bringing people together and breaking down barriers. The film also highlights the importance of forgiveness and reconciliation, and how leadership can make a real difference in people's lives.

One of the key takeaways from the film is the importance of perseverance in the face of adversity. Mandela faced immense challenges as he worked to unite South Africa, but he remained committed to his vision and never gave up. The film also emphasizes the importance of working together towards a common goal, even when it seems impossible.

Overall, ``Invictus'' is a well-made film with strong performances and an important message. However, one criticism is that the film could have delved deeper into some of the complex issues facing South Africa at the time, such as economic inequality and institutional racism. Despite this, the film is still a must-see and is deserving of a rating of 8 out of 10
\end{document}