\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}
{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}
\begin{document}
\begin{titlepage} % Suppresses displaying the page number on the title page and the subsequent page counts as page 1
		
		\raggedleft\rule{1pt}{\textheight} % Vertical line
		\hspace{0.05\textwidth} % Whitespace between the vertical line and title page text
		\parbox[b]{0.75\textwidth}
		{ % Paragraph box for holding the title page text, adjust the width to move the title page left or right on the page
			
			{\Huge\bfseries MIT World Peace University \\[0.5\baselineskip] \ Python Programming}\\[2\baselineskip] % Title
			{\large\textit{Assignment 2}}\\[4\baselineskip] % Subtitle or further description
			{\Large\textsc{Naman Soni Roll No. 10}} % Author name, lower case for consistent small caps
			
			\vspace{0.5\textheight} % Whitespace between the title block and the publisher
		}
		
\end{titlepage}
	\tableofcontents
	\pagebreak
\section{\textbf{Problem Statement}}
To find the largest of three
\section{\textbf{Aim}}
To write a python program to find the greatest of the three numbers.
\section{\textbf{Objectives}}
To learn and implement different forms of if..else statement.
\section{\textbf{Theory}}
\subsection{\textbf{\textit{Decision Making}}}
Following is the general form of a typical decision-making structure found in most of the
programming languages: \\
Python programming language assumes any non-zero and non-null values as TRUE, and
if it is either zero or null, then it is assumed as FALSE value.
Python programming language provides following types of decision-making statements.
\subsection{\textbf{\textit{Statement and Description}}}
\begin{enumerate}
	\item \textbf{if Statements:} An if statement consists of a boolean expression followed by one or more statements.
	\item \textbf{if else Statements:} An if statement can be followed by an optional else statement, which executes when the boolean expression is FALSE.
	\item \textbf{if else Statements:}If more then 2 conditions needs to be checked, then if..elif..else statement can be used.
	\item \textbf{nested if statements:} You can use one if or else if statement inside another if or else ifstatement(s).
\end{enumerate}
\section{\textbf{Platform}} Mac OS 64 bits\\
Python
\section{\textbf{Input}} 
Accept three integer numbers from the user
\section{\textbf{Output}} 
Display largest of three integer numbers accepted from the user
\section{\textbf{Conclusion}} 
Studied implementation of different forms of if...else statements using python programming.
\section{\textbf{FAQs}} 
 Describe following logical and relational operators with suitable examples.\\
Logical operators: and, or and not\\
Relational operators: <, >, <=, >=, ==,!=
